---
title: "report"
author: "vyndk"
date: "August 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load and preprocess Data
1. Load necessary packages
```{r packages, echo=TRUE}
library(tidyverse)
```
2. Read "activity.csv" data and summary it
```{r readData, echo=TRUE}
data <- read.csv("activity.csv", header = TRUE, na.strings  = "NA")
str(data)
summary(data)
```
3. Convert date column into Date type
```{r echo=TRUE}
data <- data %>% mutate(date = as.Date(date))
```
## Mean total number of steps taken per day
1. Remove NA value

```{r steps_per_day, echo=TRUE}
sub_data <- data %>% filter(!is.na(steps))
```

2. Calculate the total number of steps taken per day

```{r steps_per_day, echo=TRUE}
steps.per.day <- sub_data %>% group_by(date) %>% summarise(total.steps = sum(steps, na.rm = TRUE))
```

3. Make a histogram of the total number of steps taken each day

```{r histogram, echo=TRUE}
hist(steps.per.day$total.steps, main = "Histogram of total number of steps per day", xlab = "Total steps per day", col="red")
```

4. Calculate and report the mean and median of the total number of steps taken per day

```{r mean, echo=TRUE}
# mean
mean(steps.per.day$total.steps)
```

```{r median, echo=TRUE}
# median
median(steps.per.day$total.steps)
```

## Average daily activity pattern

1. Make a time series plot (i.e. type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r time_series, echo=TRUE}
# Calculate mean for each interval for all day
steps.per.interval <- sub_data %>% group_by(interval) %>% summarise(mean.steps = mean(steps, na.rm = TRUE))

# create a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days
plot(steps.per.interval$interval, steps.per.interval$mean.steps, type='l', 
     main="Average number of steps per interval across all days", xlab="Interval", 
     ylab="Average number of steps", col="red")
```

2. Find the 5-minute interval which contains the maximum number of steps on average across all the days in the dataset

``` {r interval, echo=TRUE}
interval <- steps.per.interval$interval[which.max(steps.per.interval$mean.steps)]
interval
```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

``` {r total.NA, echo=TRUE}
total.NA <- sum(is.na(data$steps))
total.NA
```

2. Create a new dataset that is equal to the original dataset but with the missing data filled in
   
   Filling in all of the missing values in the dataset with the mean of its 5-minute interval

``` {r fill_NA, echo=TRUE}
new.data <- data
for (i in steps.per.interval$interval) {
    new.data[new.data$interval == i & is.na(new.data$steps), ]$steps <- 
        steps.per.interval$mean.steps[steps.per.interval$interval == i]
}
head(new.data)
```

3. Make a histogram of the total number of steps taken each day for the new dataset
    
```{r steps_per_day_new, echo=TRUE}
steps.per.day.new <- new.data %>% group_by(date) %>% summarise(total.steps = sum(steps, na.rm = TRUE))
hist(steps.per.day.new$total.steps, main = "Histogram of total number of steps per day", xlab = "Total steps per day", col="blue")
```

4. Calculate and report the mean and median total number of steps taken per day

```{r mean_new, echo=TRUE}
# mean
mean(steps.per.day.new$total.steps)
```

```{r median_new, echo=TRUE}
# median
median(steps.per.day.new$total.steps)
```
    We can see that the mean and median for the new dataset are identical.
    